---
title: "Explaining analyses conducted on Scott Creek Lagoon fish stable isotope data"
subtitle: "Drought-induced sympatry of two salmonid species: Feeding ecology and trophic niches of potential competitors"
authors: "Joseph D. Kiernan and Rosealea M. Bond"
date: "2025-12-01"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    df_print: paged
---

```{r initial setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
## Summary and analysis of natural abundance stable isotope data

This document explains the analyses performed on the stable isotope samples collected in support of the manuscript titled, "Drought-induced sympatry of two salmonid species: Feeding ecology and trophic niches of potential competitors". A separate R markdown file explains the analyses conducted on fish diet (stomach content) samples. Questions regarding these analyses or the R markdown should be directed to Dr. Joseph Kiernan at joseph.kiernan@noaa.gov.


```{r Libraries, echo=FALSE, message=FALSE, warning=FALSE}

rm(list=ls()) #clear environment

# Install and load the necessary packages
library(ggplot2)
library(reshape2)
library(plyr); library(dplyr) #specific load order to avoid potential conflicts and problems
library(tidyr)
library(ggridges)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(readr)    
library(knitr)
library(kableExtra)
library(coin)
```
Data file: "Lagoon_fish_isotope.csv". 

This data set includes only age-1+ (yearling) coho salmon and steelhead. Carbon stable isotope values (header "d13C_corr") are lipid corrected values; we mathematically lipid-corrected all fin clips with a C:N ratio > 3.5 (n = 77; 15.6% of the samples). Specifically, we followed the recommendation of Skinner et al. (2016) and applied the nonlinear normalization model developed by Kiljunen et al. (2006) with the Post et al. (2007) percentage lipid calculation. 

1. Percentage Lipid Calculation (Post et al. 2007) utilizes the linear relationship between C:N and lipid content ($L$) proposed for aquatic animals:
$$L = -20.54 + (7.24 \times C:N)$$ 

2. Nonlinear Normalization Model (Kiljunen et al. 2006) takes that $L$ value and applies the modified parameters ($D = 7.018$ and $I = 0.048$) to the McConnaughey & McRoy (1979) equation to calculate the isotopic shift: 
($\Delta\delta^{13}C$):$$\Delta\delta^{13}C = 7.018 \times \left(0.048 + \frac{3.90}{1 + 287 / L}\right)$$

3. Final Correction
$$\delta^{13}C_{corr} = \delta^{13}C_{raw} + \Delta\delta^{13}C$$
This (\delta^{13}C_{corr}) is the data pipeline that is used in the analysis. See methods section of manuscript for additional details.

```{r Load full isotope data}

# Create universal input and output directories
input_dir <- "Lagoon_SIA_Input_Data"
sia_dir <- "Lagoon_SIA_Results"

if (!dir.exists(input_dir)) {
  dir.create(input_dir)
}

# Create Results Directory and Analysis Subfolders
output_dirs <- c(
  sia_dir,
  file.path(sia_dir, "SIBER"),
  file.path(sia_dir, "NicheROVER"),
  file.path(sia_dir, "Figures"),
  file.path(sia_dir, "Tables")
)

for (dir in output_dirs) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)
  }
}

if (!dir.exists(sia_dir)) {
  dir.create(sia_dir)
}

# Read the data using file.path() and the input directory
data_full <- read.csv(file.path(input_dir, "Lagoon_fish_isotope.csv"))
# The full data set contains additional fisheries metadata which do not contribute to the analysis

# To list months of the study chronologically rather than alphabetically.
data_full$Month <- factor(data_full$Month, levels =c("Jun","Jul","Aug","Sep","Oct","Nov","Dec"), ordered = TRUE)

data<-subset(data_full, select=c(Month, Species, Length_mm, d13C_corr, d15N)) #should be 494 observations of 5 variables
# The reduced data set includes five variables of interest: (1) sample month; (2) fish species code; (3) fish fork length (mm); (4) fish d13C value; and (5) fish d15N value.
```

## Summarize stable isotope data
Purpose: to summarize the stable isotope data for each salmonid species (Onki = coho salmon and Onmy = steelhead) by month.
```{r Summarize stable isotope data}
# Calculate summary metrics
isotope.fish.summary<-ddply(data, c("Month", "Species"), summarise,
          count = length(Species), #sample size
          mFL = round(mean(Length_mm),digits=1), #mean fork length in mm
          sdFL = round(sd(Length_mm),digits=1), #standard deviation fork length in mm
          mC = round(mean(d13C_corr), digits=1), #mean d13C in per mil
          sdC = round(sd(d13C_corr), digits=1),#std deviation d13C in per mil
          CR = round(max(d13C_corr)-min(d13C_corr), digits=1), #d13C range in per mil
          mN = round(mean(d15N), digits=1), #mean d15N in per mil
          sdN = round(sd(d15N), digits=1), #std deviation d15N in per mil
          NR = round(max(d15N)-min(d15N), digits=1)) #d15N range in per mil

colnames(isotope.fish.summary) <- c ("Month", "Salmonid species", "N", "Mean", "SD", "Mean", "SD", "C-Range", "Mean", "SD", "N-Range")

# Export the summary table to the Tables folder
write.csv(isotope.fish.summary, file.path(sia_dir, "Tables", "Isotope_Fish_Summary.csv"), row.names = FALSE)

# Create a kable with corrected headers and collapsed rows
isotope.fish.summary %>%
  kbl(caption = "Summary of monthly carbon and nitrogen stable isotope ratios for lagoon-rearing coho salmon (Onki) and steelhead (Onmy). These data appear, in part, in Table 4 of the manuscript") %>%
  kable_classic(font = "Cambria") %>%
  add_header_above(c(" " = 3, 
                     "Fork length (mm)" = 2, 
                     "$\\delta^{13}$C ($\\unicode{x2030}$)"= 3, 
                     "$\\delta^{15}$N ($\\unicode{x2030}$)"= 3)) %>%
  collapse_rows(columns = 1, valign = "top")
```

## Wilcoxon rank sum test contrasting stable isotope values derived for each salmonid species
Here we test for significant differences between the monthly d13C and d15N values derived for coho salmon and steelhead. It was determined during data screening that 3 of the 18 isotope groups, due to small sample sizes, violated the assumptions necessary for applying parametric t-tests. Consequently, we conservatively use the non-parametric Wilcoxon Rank Sum test to contrast monthly differences and apply a Bonferroni correction to determine significant differences.

```{r Wilcoxon Rank Sum test contrasting isotope values}
# Load required libraries
library(rstatix)
library(tidyverse) 

# Define the correct chronological order for the months
month_order <- c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

# Filter data and format
data_filtered <- data %>%
  filter(Month %in% month_order) %>%
  mutate(Month = factor(Month, levels = month_order))

data_long <- data_filtered %>%
  pivot_longer(
    cols = c("d13C_corr", "d15N"),
    names_to = "Isotope",
    values_to = "value"
  )

# Monthly tests and effect sizes
all_wilcox_tests <- data_long %>%
  group_by(Month, Isotope) %>%
  rstatix::wilcox_test(value ~ Species) 

all_wilcox_eff <- data_long %>%
  group_by(Month, Isotope) %>%
  rstatix::wilcox_effsize(value ~ Species)

all_results <- all_wilcox_tests %>%
  # Corrected join parameters here
  left_join(all_wilcox_eff, by = c("Month", "Isotope", ".y.", "group1", "group2", "n1", "n2")) %>%
  mutate(Period = as.character(Month))

# Overall tests and effect sizes
overall_wilcox_tests <- data_long %>%
  group_by(Isotope) %>%
  rstatix::wilcox_test(value ~ Species) 

overall_wilcox_eff <- data_long %>%
  group_by(Isotope) %>%
  rstatix::wilcox_effsize(value ~ Species)

overall_results <- overall_wilcox_tests %>%
  # Corrected join parameters here
  left_join(overall_wilcox_eff, by = c("Isotope", ".y.", "group1", "group2", "n1", "n2")) %>%
  mutate(Period = "Overall")

# Combine results and format for downstream code
full_results_df <- bind_rows(all_results, overall_results) %>%
  rstatix::add_significance(p.col = "p") %>% 
  select(Period, Isotope, group1, group2, statistic, p, p.signif, effsize) %>%
  mutate(effsize = round(effsize, 3))

# Write results to the SIA folder
write_csv(full_results_df, file.path(sia_dir, "Tables", "Isotope_Wilcoxon_Results.csv"))
```

### Summary table of results for Wilcoxon tests. 
These results are presented in Supplemental Table S4 in the manuscript. Note: Effect sizes are typically interpreted as small (0.1 to < 0.3), moderate (0.3 to < 0.5), and large (> 0.5).

```{r Summarizing Wilcoxon test results, results='asis'}
# This chunk uses the full_results_df created in the previous chunk
# Reshape data (Pivot wider)
wide_results <- full_results_df %>%
  pivot_wider(
    id_cols = Period, 
    names_from = Isotope,
    values_from = c(statistic, p, effsize),
    names_glue = "{Isotope}_{.value}" 
  ) %>%
  dplyr::rename(Month = Period)

# Reorder rows (chronological)
month_order <- c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Overall")
wide_results <- wide_results %>%
  mutate(Month = factor(Month, levels = month_order)) %>%
  arrange(Month)

# Format Carbon p-values
wide_results$d13C_corr_p <- cell_spec(
  format(round(wide_results$d13C_corr_p, 3), nsmall=3),
  bold = ifelse(wide_results$d13C_corr_p < 0.006, TRUE, FALSE),
  color = ifelse(wide_results$d13C_corr_p < 0.006, "#D55E00", "black")
)

# Format Nitrogen p-values
wide_results$d15N_p <- cell_spec(
  format(round(wide_results$d15N_p, 3), nsmall=3),
  bold = ifelse(wide_results$d15N_p < 0.006, TRUE, FALSE),
  color = ifelse(wide_results$d15N_p < 0.006, "#D55E00", "black")
)

# Create summary table and assign it to an object
wilcoxon_table_out <- wide_results %>%
  select(Month, 
         d13C_corr_statistic, d13C_corr_effsize, d13C_corr_p, 
         d15N_statistic, d15N_effsize, d15N_p) %>%
  kbl(
    caption = "Table S4. Summary of Wilcoxon rank-sum tests and effect sizes (r). Significant p-values after Bonferroni correction are highlighted.",
    col.names = c("Month", "W Stat", "Effect (r)", "p-value", "W Stat", "Effect (r)", "p-value"),
    digits = 3,
    escape = FALSE
  ) %>%
  kable_classic(font = "Cambria", full_width = FALSE) %>%
  add_header_above(c(" " = 1, "$\\delta^{13}$C ($\\unicode{x2030}$)" = 3, "$\\delta^{15}$N ($\\unicode{x2030}$)" = 3)) %>%
  column_spec(1, width = "2cm") %>%     
  column_spec(2:7, width = "2.0cm")

# Save the fully formatted table
save_kable(wilcoxon_table_out, file = file.path(sia_dir, "Tables", "Table_S4_Wilcoxon_Formatted.html"))

# Print the table to the document
wilcoxon_table_out
```

## Calculate total area (TA) and standard ellipse area corrected for small sample size (SEAc).

```{r Calculate and Format TA and SEAc Metrics (Monthly & Overall), echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())

# Load libraries needed for this chunk
library(SIBER)
library(dplyr)
library(tidyr)
library(tibble)

# Re-add directory paths after the environment wipe
input_dir <- "Lagoon_SIA_Input_Data"
sia_dir <- "Lagoon_SIA_Results"

if (!dir.exists(sia_dir)) {
  dir.create(sia_dir)
}

# Set seed to yield comparable results 
set.seed(8675309)

# Read the data using file.path() and the input directory
data <- read.csv(file.path(input_dir, "Lagoon_fish_isotope.csv"))

# Prepare main SIBER data
siber.data <- data[, c("d13C_corr", "d15N", "Species", "Month")]
names(siber.data) <- c("iso1", "iso2", "group", "community")
siber.data$community <- factor(siber.data$community, 
                               levels = c("Jun","Jul","Aug","Sep","Oct","Nov","Dec"))

# Create SIBER object for all months
siber.monthly.obj <- createSiberObject(siber.data)

# Calculate Maximum Likelihood (ML) group metrics for each month
group.ML.monthly <- groupMetricsML(siber.monthly.obj)

# To get "Overall" metrics, we create a copy of the data and assigned all records to a single "Overall" community.
siber.overall.data <- siber.data
siber.overall.data$community <- "Overall"

# Create a new SIBER object for the "Overall" data
siber.overall.obj <- createSiberObject(siber.overall.data)

# Calculate ML metrics for the "Overall" groups
group.ML.overall <- groupMetricsML(siber.overall.obj)

# Bind the monthly and overall matrices together
combined.metrics <- cbind(group.ML.monthly, group.ML.overall)

# Convert the matrix to a clean data.frame for the table
metrics.table <- as.data.frame(t(combined.metrics)) %>%
  rownames_to_column(var = "Group") %>%
  
  # Select only the columns needed to populate Table 4 of manuscript (i.e.,TA & SEAc)
  select(Group, TA, SEAc) %>%
  # Split the group name (e.g., "Jun.Onki") into two columns
  separate(Group, into = c("Month", "Species"), sep = "\\.") %>%
  # Ensure correct ordering of sample months
  mutate(Month = factor(Month, levels = c("Jun","Jul","Aug","Sep",
                                          "Oct","Nov","Dec","Overall"))) %>%
  arrange(Month, Species)

# Save this table to the Lagoon_SIA_Results directory:
write.csv(metrics.table, file.path(sia_dir, "SIBER", "SIBER_Metrics_Summary.csv"), row.names = FALSE)

metrics.wide <- metrics.table %>%
  # Select columns needed
  select(Month, Species, TA, SEAc) %>%
  # Pivot wider to get TA and SEAc columns for each species
  pivot_wider(
    names_from = Species,
    values_from = c(TA, SEAc)
  ) %>%
  # Reorder columns to [Month, TA_Onki, SEAc_Onki, TA_Onmy, SEAc_Onmy]
  select(Month, TA_Onki, SEAc_Onki, TA_Onmy, SEAc_Onmy)

# Define column names for the inner row
col_names <- c("Month", "TA", "SEAc", "TA", "SEAc")
```

### Create summary table of TA and SEAc metrics
These values contribute, in part, to table 4 in the manuscript

```{r Create table of TA and SEAc Metrics (Monthly & Overall)}
# Create the final kable output and assign it to an object
TA_SEAc_table_out <- metrics.wide %>%
  knitr::kable(
    caption = "Table 4 (in part). Monthly total area and SEAc values for each salmonid species", 
    digits = 1,
    col.names = col_names #Apply the inner column names
  ) %>%
  kable_classic(font = "Cambria", full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Coho Salmon" = 2, "Steelhead" = 2)) %>% 
  column_spec(1:5, width = "2cm")

# Save the fully formatted table to the Tables folder
save_kable(TA_SEAc_table_out, file = file.path(sia_dir, "Tables", "Table_4_TA_SEAc_Formatted.html"))

# Print the table to the document
TA_SEAc_table_out
```

### Generating stable isotope biplots for each time period. 
These data are presented as Figure 2 in the manuscript. 

```{r Figure 2, Generate isotope biplot, fig.width=6.5, fig.height=3.97}
library(grid) 
library(patchwork)

# Define labels
plot_labels <- c("Jun"="(a) Jun", "Jul"="(b) Jul", "Aug"="(c) Aug", "Sep"="(d) Sep", 
                 "Oct"="(e) Oct", "Nov"="(f) Nov", "Dec"="(g) Dec", "Overall"="(h) Overall")

# Plotting function (Base plot)
create_siber_biplot <- function(plot_data, plot_code) {
  x_lims <- c(-32, -17)
  y_lims <- c(8, 15)
  
  ggplot(plot_data, aes(x = iso1, y = iso2, col = group, fill = group)) +
    geom_point(size = 0.8) +
    stat_ellipse(geom = "polygon", alpha = 0.30, level = 0.40) +
    stat_ellipse(geom = "polygon", alpha = 0.0, level = 0.95) +
    scale_color_manual(values = c("Onki" = "#D55E00", "Onmy" = "#0072B2")) +
    scale_fill_manual(values = c("Onki" = "#D55E00", "Onmy" = "#0072B2")) +
    coord_cartesian(xlim = x_lims, ylim = y_lims) +
    labs(title = plot_labels[[plot_code]]) +
    theme_classic() +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(family = "serif", face = "plain", size = 10, hjust = 0),
      axis.title = element_blank(), 
      axis.text = element_text(family = "serif", size = 9, color = "black"),
      axis.line = element_line(color = "black", linewidth = 0.3), # reduced line weight
      axis.ticks = element_line(color = "black", linewidth = 0.3)
    )
}

# Data prep
monthly_data_list <- split(siber.data, siber.data$community)
ordered_keys <- c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
plot_data_list <- monthly_data_list[ordered_keys]
plot_data_list$Overall <- siber.data

# Generate base plots
all_siber_plots <- imap(plot_data_list, create_siber_biplot)

# Apply conditional formatting (ticks & legend)
final_plot_list <- map2(all_siber_plots, names(all_siber_plots), function(p, name) {
  # Top Row: Jun, Jul, Aug, Sep -> Remove x-axis text
  if (name %in% c("Jun", "Jul", "Aug", "Sep")) {
    p <- p + theme(axis.text.x = element_blank())
  }
  
  # Not First Column: Jul, Aug, Sep, Nov, Dec, Overall -> Remove Y axis text
  if (!name %in% c("Jun", "Oct")) {
    p <- p + theme(axis.text.y = element_blank())
  }
  
  # Legend relocation and resizing
  if (name == "Jun") {
    p <- p + theme(
      # Shifted slightly right and down to prevent y-axis overlap
      legend.position = c(0.28, 0.78), 
      legend.justification = c(0.5, 0.5),
      legend.background = element_rect(fill = "white", color = "black", linewidth = 0.3),
      legend.margin = margin(2, 2, 2, 2), # Tighter margins around the legend box
      legend.key.size = unit(0.7, "lines"), # Shrinks the colored boxes
      legend.text = element_text(family = "serif", size = 7), # Shrinks the legend text
      legend.title = element_blank(), 
      text = element_text(family = "serif")
    )
  } else {
    # Hide legend for all other panels
    p <- p + theme(legend.position = "none")
  }
  
  return(p)
})

# Main Grid
main_grid <- wrap_plots(final_plot_list, ncol = 4, nrow = 2) & 
  theme(text = element_text(family = "serif"))

# Global axes
y_expr <- expression(paste(delta^{15}, "N (\u2030)"))
x_expr <- expression(paste(delta^{13}, "C (\u2030)"))

# Set global axis font size to 10
y_grob <- textGrob(y_expr, rot = 90, gp = gpar(fontfamily = "serif", fontsize = 10))
x_grob <- textGrob(x_expr, gp = gpar(fontfamily = "serif", fontsize = 10))

# Assemble all
final_siber_panel <- (wrap_elements(y_grob) + main_grid + plot_layout(widths = c(1, 25))) / 
                     wrap_elements(x_grob) + 
                     plot_layout(heights = c(25, 1.1))

final_siber_panel

ggsave(file.path(sia_dir, "Figures", "Figure_2.pdf"), 
       width = 6.5, height = 3.97, units = "in", dpi = 800, 
       device = function(file, ...) quartz(file = file, type = "pdf", ...))
ggsave(file.path(sia_dir, "Figures", "Figure_2.tif"), 
       width = 6.5, height = 3.97, units = "in", dpi = 800, 
       device = "tiff", compression = "lzw")
```

Figure 2. Caption:
Stable isotope biplots (δ13C versus δ15N) for juvenile coho salmon (*Oncorhynchus kisutch*; Onki) and steelhead (*O. mykiss*; Onmy), captured in the Scott Creek lagoon during each month of the study (Jun–Dec, panels a–g respectively). Panel h shows composited data from all months of the study. Points are individual fish samples and filled ellipses are the standard ellipse areas (corrected for small sample size; SEAc). Standard ellipse areas encompass approximately 40% of the isotopic data and represent the core isotopic niche for each species. The larger ellipses contain approximately 95% of the isotopic data for each species.


## Calculating Bayesian Standard Ellipse Area (SEAb)

The previous calculations yielded point estimates (i.e., single value) of overlap based on the maximum likelihood estimated SEAc. To assess how reliable these data are, we need an associated measure of confidence/variance. This is achieved by calculating the Bayesian standard ellipse area (SEAb) and generating credible intervals.

```{r Calculate SEAb with uninformed priors, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}

set.seed(2112)

# Run the SIBER model, default settings, uninformed priors
# SIBER mcmc model parameters
parms <- list()
parms$n.iter <- 2 * 10^4 # number of iterations to run the model
parms$n.burnin <- 1 * 10^3 # discard the first set of values
parms$n.thin <- 10 # thin the posterior by this many
parms$n.chains <- 2 # run this many chains

# SIBER mcmc model priors
priors <- list()
priors$R <- 1 * diag(2)
priors$k <- 2
priors$tau.mu <- 1.0E-3

# Run SIBER mcmc
ellipses.posterior <- siberMVN(siber.monthly.obj, parms, priors)

SEA.B <- siberEllipses(ellipses.posterior)
```

Here is a preliminary look at the distribution of ellipses for each population as a SIBER density plot. The gradations denote 95, 75 and 50% credible intervals. This plot does not include the overall (study-wide) pooled estimate which is generated later. This plot, in part, will be recreated later using ggplot to produce Figure 3 in the manuscript.

```{r Creating standard SIBER density plot, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}
# Save current settings
op <- par(no.readonly = TRUE)

# Set the global axis text size before the plot function
par(cex.axis = 0.7) 

# SEA.B <- siberEllipses(ellipses.posterior)
siberDensityPlot(
                SEA.B, 
                xlab = ("Monthly contrasts (C = coho, S = steelhead, Number = month)"),
                ylab = expression("Dietary niche breadth "(SEA[b],'\u2030' ^ 2) ),
                xticklabels = c("C6", "S6",
                                "C7", "S7",
                                "C8", "S8",
                                "C9", "S9",
                                "C10", "S10",
                                "C11", "S11",
                                "C12", "S12"),
                yticklabels = NULL,
                probs = c(50, 75, 95),
                bty = "L",
                las = 1,
                ylim = c(0,5),
                main = ""
                )

# Restore original settings
par(op)

# The nfollowing line will add red x's for the ML estimated SEAc, if desired
#points(1:ncol(SEA.B), group.ML.monthly[3,], col="red", pch = "x", lwd = 2)
```

### SEAb Credible Intervals

The following section of code extracts the mode and credibility intervals (SEAb) for each group and time period. Note, with an uninformative prior, the maximum likelihood estimate should correspond to the mode of the posterior distribution.

```{r Determining SEA.B credible intervals, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}
# Examining the SEA.B credible intervals 
cr.p <- c(0.50, 0.75, 0.95) # Vector of quantiles

colnames(SEA.B) <- c('Jun.Onki', 'Jun.Onmy',
                      'Jul.Onki', 'Jul.Onmy',
                      'Aug.Onki', 'Aug.Onmy',
                      'Sep.Onki', 'Sep.Onmy',
                      'Oct.Onki', 'Oct.Onmy',
                      'Nov.Onki', 'Nov.Onmy',
                      'Dec.Onki', 'Dec.Onmy'
                      ) #re-name columns in new list

# Calculate modes and credible intervals of SEA.B
SEA.B.modes <- lapply(
  as.data.frame(SEA.B), 
  function(x,...){tmp <- hdrcde::hdr(x)$mode},
  prob = cr.p, all.modes = T) #vectors of the modes for each group's SEA.b
#SEA.B.modes

# Call to hdrcde:hdr using lapply()
SEA.B.credibles <- lapply(
  as.data.frame(SEA.B), 
  function(x,...){tmp <- hdrcde::hdr(x)$hdr},
  prob = cr.p) #vectors of the credible intervals for each group's SEA.b
#SEA.B.credibles

# Save this output to the SIBER_Output directory
write.table(SEA.B.modes, file = file.path(sia_dir, "SIBER", "SEAb_Modes.csv"), 
            append = F, sep = ',', row.names = T, col.names = NA)

write.table(SEA.B.credibles, file = file.path(sia_dir, "SIBER", "SEAb_Credible_Intervals.csv"), 
            append = F, sep = ',', row.names = T, col.names = NA)
```

Although SIBER has a built-in SEAb plot (Siber.density.plot, created above), we will prepare the SIBER output for ggplot which is easier to customize and manipulate.

```{r Draw SEA.B posteriors, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}

# Draw posteriors for SIBER ellipse area for each "species.month" combination
SEA.B.2 <- SEA.B 
colnames(SEA.B.2) <- colnames(group.ML.monthly) 
SEA.B.2 <- as.data.frame(SEA.B.2) 
SEA.B.2 <- melt(SEA.B.2) 
SEA.B.2$combo <- SEA.B.2$variable 
SEA.B.2 <- separate(data = SEA.B.2,
                    col = variable,
                    into = c("Month","Species"),
                    sep = "\\.",
                    extra = "merge")
SEA.B.2$Month <- factor(SEA.B.2$Month,
        levels = c('Jun','Jul','Aug','Sep','Oct','Nov','Dec')) #re-order months chronologically
```
Note: the warning "No id variables; using all as measure variables" is the precise structure we want for the data

Next we re-run the analysis for to generate the "overall" (all monthly data combined) estimates foe each salmonid spp.

```{r Generate "Overall" SIBER model, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}

#"Overall" SIBER model, default settings, uninformed priors

set.seed(2112)

# SIBER mcmc model parameters
parms <- list()
parms$n.iter <- 2 * 10^4 #number of iterations to run the model
parms$n.burnin <- 1 * 10^3 #discard the first set of values
parms$n.thin <- 10 #thin the posterior by this many
parms$n.chains <- 2 #run this many chains

# SIBER mcmc model priors
priors <- list()
priors$R <- 1 * diag(2)
priors$k <- 2
priors$tau.mu <- 1.0E-3

# Run SIBER mcmc
ellipses.posterior.overall <- siberMVN(siber.overall.obj, parms, priors) 

# Calculate SEA.B for the "overall" data
SEA.B.overall <- siberEllipses(ellipses.posterior.overall) 

# Rename the columns of the "overall" results using the ML object names
# Note: this assumes group.ML.overall was calculated earlier
overall_group_names <- colnames(group.ML.overall) 
colnames(SEA.B.overall) <- overall_group_names
```

```{r Combine Monthly and Overall SEA.B results}
# SEA.B.overall is the combined result from the chunk above.

# Combine the monthly and overall SEA.B results into a single object
SEA.B.full <- cbind(SEA.B, SEA.B.overall)

# Optional code: can reassign the column names of the full matrix for a final check 
full_group_names <- c(colnames(SEA.B), colnames(SEA.B.overall))
colnames(SEA.B.full) <- full_group_names

# Additional data wrangling steps, using the new combined object
SEA.B.2 <- SEA.B.full # Use the full combined object, which includes 'Overall' data
SEA.B.2 <- as.data.frame(SEA.B.2) 
SEA.B.2 <- melt(SEA.B.2)
SEA.B.2$combo <- SEA.B.2$variable
SEA.B.3 <- SEA.B.2

# The remaining code processes the 'Overall' data:
SEA.B.3 <- separate(data = SEA.B.3, col = variable, into = c("Month","Species"), sep = "\\.", extra = "merge") 
SEA.B.3$Month <- factor(SEA.B.3$Month, levels = c('Jun','Jul','Aug','Sep','Oct','Nov','Dec','Overall')) #re-order months 

```
Again, "No id variables; using all as measure variables" is the precise structure we want for the data.

```{r Creating the "Overall" SIBER density plot, echo=TRUE, message=TRUE, warning=TRUE, cache=TRUE}

siberDensityPlot(
               SEA.B.overall, 
                xlab = c("Overall"), 
                 ylab = expression("Dietary niche breadth " (SEA[b],'\u2030' ^2) ),
                 xticklabels = c("Coho salmon", "Steelhead"),
                 yticklabels = NULL,
                 probs = c(50, 75, 95),
                 bty = "L",
                 las = 1,
                 ylim = c(0,8),
                 main = ""
                 )

#The following line of code (currently suppressed) will add red x's for the ML estimated SEAc if desired
#points(1:ncol(SEA.B), group.ML.overall[3,], col="red", pch = "x", lwd = 2)
```


```{r Examining the "Overall" SEA.B credible intervals}
# Examining the SEA.B credible intervals 
cr.p <- c(0.50, 0.75, 0.95) # vector of quantiles

# Calculate modes and credible intervals of SEA.B
SEA.B.overall.modes <- lapply(
  as.data.frame(SEA.B.overall), 
  function(x,...){tmp <- hdrcde::hdr(x)$mode},
  prob = cr.p, all.modes = T) #vectors of the modes for each group's SEA.b
#SEA.B.overall.modes

# Call to hdrcde:hdr using lapply()
SEA.B.overall.credibles <- lapply(
  as.data.frame(SEA.B.overall), 
  function(x,...){tmp <- hdrcde::hdr(x)$hdr},
  prob = cr.p) #vectors of the credible intervals for each group's SEA.b
#SEA.B.overall.credibles

# Save this output to the SIBER_Output directory
write.table(SEA.B.overall.modes, file = file.path(sia_dir, "SIBER", "SEAb_Overall_Modes.csv"), 
            append = F, sep = ',', row.names = T, col.names = NA)

write.table(SEA.B.overall.credibles, file = file.path(sia_dir, "SIBER", "SEAb_Overall_Credible_Intervals.csv"), 
            append = F, sep = ',', row.names = T, col.names = NA)
```

Here we combine SEA.B (monthly) and SEA.B.overall into SEA.B.full

```{r Create SEA.B.full which combines monthly and overall estimates, echo=TRUE, message=TRUE, warning=TRUE}

# Combination of monthly (SEA.B) and overall (SEA.B.overall) results into a full matrix
SEA.B.full <- cbind(SEA.B, SEA.B.overall)

# Define the *exact* column names for the combined matrix
full_col_names <- c('Jun.Onki', 'Jun.Onmy',
                    'Jul.Onki', 'Jul.Onmy',
                    'Aug.Onki', 'Aug.Onmy',
                    'Sep.Onki', 'Sep.Onmy',
                    'Oct.Onki', 'Oct.Onmy',
                    'Nov.Onki', 'Nov.Onmy',
                    'Dec.Onki', 'Dec.Onmy',
                    'Overall.Onki', 'Overall.Onmy')
colnames(SEA.B.full) <- full_col_names

#Melt the data for ggplot (creating SEA.B.3)
SEA.B.3 <- as.data.frame(SEA.B.full) %>% 
  #Melt the wide data into long format
  melt(variable.name = "combo", value.name = "value") %>%
  #Split the 'combo' column into Month and Species
  separate(col = combo, into = c("Month", "Species"), sep = "\\.") %>%
  #Ensure chronological and 'Overall' order for Month factor
  mutate(Month = factor(Month, 
                        levels = c('Jun','Jul','Aug','Sep','Oct','Nov','Dec','Overall'),
                        ordered = TRUE))
```

```{r Probability of different SEA.B}

set.seed(123)

P_jun<- sum( SEA.B[,1] < SEA.B[,2] ) / nrow(SEA.B)
P_jul<- sum( SEA.B[,3] < SEA.B[,4] ) / nrow(SEA.B)
P_aug<- sum( SEA.B[,5] < SEA.B[,6] ) / nrow(SEA.B)
P_sep<- sum( SEA.B[,7] < SEA.B[,8] ) / nrow(SEA.B)
P_oct<- sum( SEA.B[,9] < SEA.B[,10] ) / nrow(SEA.B)
P_nov<- sum( SEA.B[,11] < SEA.B[,12] ) / nrow(SEA.B)
P_dec<- sum( SEA.B[,13] < SEA.B[,14] ) / nrow(SEA.B)
P_overall<- sum(SEA.B.overall[,1] < SEA.B.overall[,2] ) / nrow(SEA.B.overall)

prob_summary_df <- data.frame(
  Month = c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Overall"),
  Probability = c(P_jun, P_jul, P_aug, P_sep, P_oct, P_nov, P_dec, P_overall)
)

prob_summary_df$Month <- factor(prob_summary_df$Month, 
                                levels = c("Jun","Jul","Aug","Sep","Oct","Nov","Dec","Overall"),
                                ordered = TRUE)

prob_summary_df %>%
  kbl(
    caption = "Probability that the isotopic niche breadth (SEAb) of coho salmon (Onki) is less than steelhead (Onmy)",
    digits = 2,
    col.names = c("Month", "Probability")
  ) %>%
  kable_classic(font = "Cambria", full_width = FALSE) %>%

  add_header_above(c(" " = 1, "Pr(SEAb Onki < SEAb Onmy)" = 1))
```

The above data are not included in the manuscript but confirm that the isotopic niche breadth of lagoon-rearing steelhead (Onmy) was generally greater than that of coho salmon (Onki).

### Determining monthly and overall overlap in SIBER

Note: this chunk of code is very slow to run.
```{r Determining monthly and overall overlap in SIBER, cache=TRUE}

set.seed(2564)
#Define the ellipses:

ellipse1 = "Jun.Onki"
ellipse2 = "Jun.Onmy"
ellipse3 = "Jul.Onki"
ellipse4 = "Jul.Onmy"
ellipse5 = "Aug.Onki"
ellipse6 = "Aug.Onmy"
ellipse7 = "Sep.Onki"
ellipse8 = "Sep.Onmy"
ellipse9 = "Oct.Onki"
ellipse10 = "Oct.Onmy"
ellipse11 = "Nov.Onki"
ellipse12 = "Nov.Onmy"
ellipse13 = "Dec.Onki"
ellipse14 = "Dec.Onmy"
ellipse15 = "Overall.Onki"
ellipse16 = "Overall.Onmy"

jun.bayes95.overlap <- bayesianOverlap(ellipse1, ellipse2, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
jul.bayes95.overlap <- bayesianOverlap(ellipse3, ellipse4, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
aug.bayes95.overlap <- bayesianOverlap(ellipse5, ellipse6, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
sep.bayes95.overlap <- bayesianOverlap(ellipse7, ellipse8, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
oct.bayes95.overlap <- bayesianOverlap(ellipse9, ellipse10, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
nov.bayes95.overlap <- bayesianOverlap(ellipse11, ellipse12, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
dec.bayes95.overlap <- bayesianOverlap(ellipse13, ellipse14, ellipses.posterior,
                                   draws = 1000, p.interval = 0.95, n = 100)
overall.bayes95.overlap <- bayesianOverlap(ellipse15, ellipse16, ellipses.posterior.overall,
                                   draws = 1000, p.interval = 0.95, n = 100)
```

The next chunk calculates the distribution of overlap based on the posterior distributions of the fitted ellipses. Overlap is expressed as a proportion of the non-overlapping area of the two ellipses so values should range from 0 to 1.

```{r Distribution of 95 overlap}
jun.bayes.prop.95.over <- (jun.bayes95.overlap[,3] / (jun.bayes95.overlap[,2] + 
                                                jun.bayes95.overlap[,1] -
                                                jun.bayes95.overlap[,3]))
jul.bayes.prop.95.over <- (jul.bayes95.overlap[,3] / (jul.bayes95.overlap[,2] + 
                                                jul.bayes95.overlap[,1] -
                                                jul.bayes95.overlap[,3]))
aug.bayes.prop.95.over <- (aug.bayes95.overlap[,3] / (aug.bayes95.overlap[,2] + 
                                                aug.bayes95.overlap[,1] -
                                                aug.bayes95.overlap[,3]))
sep.bayes.prop.95.over <- (sep.bayes95.overlap[,3] / (sep.bayes95.overlap[,2] + 
                                                sep.bayes95.overlap[,1] -
                                                sep.bayes95.overlap[,3]))
oct.bayes.prop.95.over <- (oct.bayes95.overlap[,3] / (oct.bayes95.overlap[,2] + 
                                                oct.bayes95.overlap[,1] -
                                                oct.bayes95.overlap[,3]))
nov.bayes.prop.95.over <- (nov.bayes95.overlap[,3] / (nov.bayes95.overlap[,2] + 
                                                nov.bayes95.overlap[,1] -
                                                nov.bayes95.overlap[,3]))
dec.bayes.prop.95.over <- (dec.bayes95.overlap[,3] / (dec.bayes95.overlap[,2] + 
                                                dec.bayes95.overlap[,1] -
                                                dec.bayes95.overlap[,3]))
overall.bayes.prop.95.over <- (overall.bayes95.overlap[,3] / (overall.bayes95.overlap[,2] + 
                                                overall.bayes95.overlap[,1] -
                                                overall.bayes95.overlap[,3]))
#jun.bayes.prop.95.over
#jul.bayes.prop.95.over
#aug.bayes.prop.95.over
#sep.bayes.prop.95.over
#oct.bayes.prop.95.over
#nov.bayes.prop.95.over
#dec.bayes.prop.95.over
#overall.bayes.prop.95.over
```

The next chunk creates a new data.frame (name = bayes.95.overlap.plotting) that combines the above posterior distributions (of proportional overlap) for data summary and visualization.

```{r Overlap based on the posterior distributions of SEAB}

bayes.95.overlap.plotting<-data.frame(jun.bayes.prop.95.over,
                                      jul.bayes.prop.95.over,
                                      aug.bayes.prop.95.over,
                                      sep.bayes.prop.95.over,
                                      oct.bayes.prop.95.over,
                                      nov.bayes.prop.95.over,
                                      dec.bayes.prop.95.over,
                                      overall.bayes.prop.95.over
                                      )
colnames(bayes.95.overlap.plotting)<-c("Jun","Jul","Aug","Sep","Oct","Nov","Dec","Overall")

#head(bayes.95.overlap.plotting)

boxplot(as.data.frame(bayes.95.overlap.plotting),
        main = "Overlap (95% CI) based on the posterior distributions of SEAb", 
        xlab ="Month",
        ylab = "Proportion niche overlap",
        las =1,
        ylim = c(0, 1.0),
        boxwex = 0.4
)
```
Overlap values are presented as a proportion of the total area that overlaps in the two 95% ellipses and overlap values range from 0 when the ellipses are completely separated to 1.0 when the two ellipses fully overlap.

The above plot is not included in the manuscript but is useful to explore the data. The next chunk of code extracts summary statistics for each distribution.

### Mean, median and modal overlap values based on the posterior distributions of SEAB

```{r Mean, median and mode overlap values based on the posterior distributions of SEAB}

df <- as.data.frame(bayes.95.overlap.plotting)

# Calculate the mean for each column (Month and Overall)
all_means <- sapply(df, mean, na.rm = TRUE)

# Calculate the median for each column (Month and Overall)
all_medians <- sapply(df, median, na.rm = TRUE)

# Find the peak of the density (the "mode") for each column (Month and Overall)
get_density_peak <- function(x, na.rm = TRUE) {
  if (na.rm) {
    x <- na.omit(x)
  }
  if (length(x) == 0) {
    return(NA)
  }
# Create density estimate
  d <- density(x)
# Return the 'x' value that has the highest 'y' value (the peak of ther distribution)
  return(d$x[which.max(d$y)])
}

# Apply this function to every column
all_modes <- sapply(df, get_density_peak)

# Create the final table
results_table <- data.frame(
  Mean = all_means,
  Median = all_medians,
  Mode = all_modes
)

# Set the row names to be the column names (Months and Overall)
rownames(results_table) <- names(df)

library(tibble) #for rownames_to_column

# Convert row names to a column and store it as a clean data frame
final_overlap_stats <- results_table %>%
  rownames_to_column(var = "Month")

# Export the clean data frame to your centralized output folder
write.csv(final_overlap_stats, file.path(sia_dir, "SIBER", "SEAb_Posterior_Overlap_Stats.csv"), row.names = FALSE)

# Generate the HTML table for the document
final_overlap_stats %>%
  kbl(
    caption = "Summary statistics for the proportional niche overlap (SEAb, 95% CrI) between coho salmon and steelhead generated by SIBER.",
    digits = 2,
    col.names = c("Month", "Mean", "Median", "Mode")
  ) %>%
  kable_classic(font = "Cambria", full_width = TRUE)

```

### Creating Figure 3, Isotopic Diet Breadth

Here we make an alternative figure to the SIBER DENSITY PLOT that illustrates the mean SEAb value and credible intervals for each posterior distribution.

```{r Figure 3, echo=TRUE, message=FALSE, warning=FALSE}

library(patchwork)
library(grid)
library(tidybayes)
library(dplyr) 

# Pre-calculate the credible intervals
ci_95 <- SEA.B.3 %>% group_by(Month, Species) %>% mean_qi(value, .width = 0.95)
ci_75 <- SEA.B.3 %>% group_by(Month, Species) %>% mean_qi(value, .width = 0.75)
ci_50 <- SEA.B.3 %>% group_by(Month, Species) %>% mean_qi(value, .width = 0.50)

plot_labels <- c("Jun"="(a) Jun", "Jul"="(b) Jul", "Aug"="(c) Aug", "Sep"="(d) Sep", 
                 "Oct"="(e) Oct", "Nov"="(f) Nov", "Dec"="(g) Dec", "Overall"="(h) Overall")
ordered_keys <- c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Overall")

# Function to build individual panels (this forces L-shapes on every plot)
create_diet_plot <- function(month_key) {
  
  d95 <- ci_95 %>% filter(Month == month_key)
  d75 <- ci_75 %>% filter(Month == month_key)
  d50 <- ci_50 %>% filter(Month == month_key)
  
  p <- ggplot() +
    geom_crossbar(data = d95, aes(x = Species, y = value, ymin = .lower, ymax = .upper, fill = Species),
                  width = 0.15, color = "black", fatten = 0, linewidth = 0.3) +
    geom_crossbar(data = d75, aes(x = Species, y = value, ymin = .lower, ymax = .upper, fill = Species),
                  width = 0.25, color = "black", fatten = 0, linewidth = 0.3) +
    geom_crossbar(data = d50, aes(x = Species, y = value, ymin = .lower, ymax = .upper, fill = Species),
                  width = 0.50, color = "black", fatten = 0, linewidth = 0.3) +
    geom_point(data = d50, aes(x = Species, y = value), size = 1.2, color = "black") +
    
    scale_fill_manual(values = c("Onki" = "#EAAF80", "Onmy" = "#80B9D9")) +
    scale_y_continuous(limits = c(0, 7.3), breaks = seq(0, 8, 1)) +
    
    # Place panel tag inside the top-left corner
    annotate("text", x = -Inf, y = Inf, label = plot_labels[[month_key]], 
             hjust = -0.10, vjust = 1.5, family = "serif", size = 2.8) +
    
    theme_classic() +
    theme(
      axis.title = element_blank(), # Removed local titles; we will use global ones
      axis.text.y = element_text(family = "serif", size = 9, color = "black"),
      axis.text.x = element_text(family = "serif", size = 7.0, color = "black"),
      axis.line = element_line(color = "black", linewidth = 0.3),# reduced weight of axis lines
      axis.ticks = element_line(color = "black", linewidth = 0.3),
      axis.ticks.length = unit(0.1, "cm"),
      
      legend.position = "none"
    )
  # Remove Y-axis numbers for all panels except the first one (Jun)
  if (month_key != "Jun") {
    p <- p + theme(axis.text.y = element_blank(),
                   plot.margin = margin(t = 5, r = 1, b = 5, l = 1, unit = "pt")
    )
  }
  
  return(p)
}

# Generate the list of plots and stitch them together in a 1x8 grid
all_diet_plots <- lapply(ordered_keys, create_diet_plot)
main_grid <- wrap_plots(all_diet_plots, ncol = 8, nrow = 1)

# Create global axes
y_expr <- expression(paste("Dietary niche breadth (", SEA[b], ", \u2030"^2, ")"))
y_grob <- textGrob(y_expr, rot = 90, gp = gpar(fontfamily = "serif", fontsize = 10))
x_grob <- textGrob("Salmonid species", gp = gpar(fontfamily = "serif", fontsize = 10))

# Assemble the final panel with patchwork
Diet.breadth.plot <- (wrap_elements(y_grob) + main_grid + plot_layout(widths = c(0.75, 25))) / 
                     wrap_elements(x_grob) + 
                     plot_layout(heights = c(25, 1.0))

Diet.breadth.plot

ggsave(file.path(sia_dir, "Figures", "Figure_3.pdf"), 
       width = 6.5, height = 3.97, units = "in", dpi = 800, 
       device = function(file, ...) quartz(file = file, type = "pdf", ...))
ggsave(file.path(sia_dir, "Figures", "Figure_3.tif"), 
       width = 6.5, height = 3.97, units = "in", dpi = 800, 
       device = "tiff", compression = "lzw")
```
Figure 3 legend: 
Estimates of dietary niche breadth (‰^2) for juvenile coho salmon (*Oncorhynchus kisutch*; Onki) and steelhead (*O. mykiss*; Onmy) captured in the Scott Creek lagoon. Species-specific estimates were generated for each month of the study (panels a–g) and an overall estimate (h) was produced from pooled data (all months combined) using posterior estimates of Bayesian standard ellipse areas (SEAb). Black dots represent the mean SEAb value after 10,000 iterations and box widths from wide to narrow represent 50, 75, and 95% credible intervals. 

```{r, quick QC for data validation}
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

# Re-use the density peak function from earlier in the script to calculate the Mode
get_density_peak <- function(x) {
  d <- density(x)
  return(d$x[which.max(d$y)])
}

# Extract all metrics from the SEA.B.3 posterior distributions
verification_table <- SEA.B.3 %>%
  group_by(Month, Species) %>%
  summarise(
    Mean = mean(value),
    Median = median(value),
    Mode = get_density_peak(value),
    # 50% CrI bounds (25th to 75th percentiles)
    CrI_50_low = quantile(value, 0.25),
    CrI_50_up = quantile(value, 0.75),
    # 75% CrI bounds (12.5th to 87.5th percentiles)
    CrI_75_low = quantile(value, 0.125),
    CrI_75_up = quantile(value, 0.875),
    # 95% CrI bounds (2.5th to 97.5th percentiles)
    CrI_95_low = quantile(value, 0.025),
    CrI_95_up = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  # Format the numbers to 2 decimal places
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Export the table to the SIA folder
write.csv(verification_table, file.path(sia_dir, "SIBER", "QC_Data_Validation.csv"), row.names = FALSE)

# Print the verification table
verification_table %>%
  kbl(
    caption = "SEAb Posterior Distribution Central Tendencies and Credible Intervals",
    align = "llrrrrrrrrr"
  ) %>%
  kable_classic(font = "Cambria", full_width = FALSE) %>%
  add_header_above(c(" " = 2, "Central Tendencies" = 3, "50% CrI" = 2, "75% CrI" = 2, "95% CrI" = 2))
```

##NicheROVER - Determining isotopic niche overlap
          
```{r nicheROVER setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This next section uses the R package “nicheROVER” to estimate dietary niche overlap between lagoon-rearing coho salmon and steelhead during each month of the study (Jun-Dec, 2015). We will run the model separately for each time period (month), and perform one run on all data pooled to create an "overall" estimate. 

```{r Global nicheRover libraries and parameters}

# Clean environment and set seed
rm(list=ls())
graphics.off()
set.seed(2112)

# Load required libraries
library(nicheROVER)
library(ggplot2)
library(ggeasy) 
library(patchwork)  
library(kableExtra) 

# Re-add directory paths after the environment wipe
input_dir <- "Lagoon_SIA_Input_Data"
sia_dir <- "Lagoon_SIA_Results"
if (!dir.exists(sia_dir)) {
  dir.create(sia_dir)
}

summary_stats_list <- list() #To store table data

# Read the data
tryCatch({
  Fish.isotope.data <- read.csv(file.path(input_dir, "lagoon_fish_isotope.csv"), header = T)
}, error = function(e) {
  stop("Failed to read 'lagoon_fish_isotope.csv'. Make sure the file is in the 'Lagoon_SIA_Input_Data' directory.")
})

# Define analysis time periods
periods <- c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Overall")

# Initialize lists to store results
# These store the ggplot objects and data frames for each sample period
plot_list <- list()
all_plot_data_list <- list()

# Define constants
clrs <- c("#D55E00", "#0072B2") # Onki = Vermilion, Onmy = Blue
nsamples <- 1000 #Number of posterior samples

```

The following chunk loops through each period, subsets the data, runs the `nicheROVER` analysis, saves all CSV and plot outputs, and stores the `ggplot` object for the final panel (figure 4 in the manuscript).

```{r Main nicheROVER analysis loop, warning=FALSE, message=FALSE, cache=TRUE}

for (period in periods) {
  
  cat(paste("\n--- Processing:", period, "---\n"))
  
# Subset and format data
  nicheROVER.df <- NULL #Ensure clean data frame
  
  if (period == "Overall") {
# For "Overall", subset only by species
    nicheROVER.df <- subset(Fish.isotope.data, Species %in% c("Onki", "Onmy"))
  } else {
# For specific months, subset by Month and Species
    nicheROVER.df <- subset(Fish.isotope.data, Month %in% period & Species %in% c("Onki", "Onmy"))
  }
  
# Reorder and rename columns for nicheROVER
  nicheROVER.df <- nicheROVER.df[c("Species", "d13C_corr", "d15N")]
  names(nicheROVER.df) <- c("species", "D13C", "D15N")
  
# Check data sufficiency
  species_counts <- table(nicheROVER.df$species)
  
# Check that we have both species and at least 3 samples for each (niw.post needs n > p)
  if (length(species_counts) < 2 || any(species_counts < 3)) {
    cat(paste("Skipping", period, "- insufficient data (n < 3) for one or both species.\n"))
    print(species_counts)
    next #Skip to the next iteration
  }
  
  cat("Sample sizes:\n")
  print(species_counts)
  
### Run nicheROVER analysis###
  
# Aggregate data for mean value check
  cat("Mean isotope values:\n")
  print(aggregate(nicheROVER.df[2:3], nicheROVER.df[1], mean))
  
# Create fish.par object (posterior draws)
  fish.par <- tapply(1:nrow(nicheROVER.df), nicheROVER.df$species,
                     function(ii) niw.post(nsamples = nsamples, X = nicheROVER.df[ii, 2:3]))
  
### Niche Overlap Calculation
# Calculate overlap for both 95% and 99%
  over.stat <- overlap(fish.par, nreps = nsamples, nprob = 1e4, alpha = c(.95, 0.99))
  
# Calculate and save mean overlap
  over.mean <- apply(over.stat, c(1:2, 4), mean) * 100
  cat("Mean Overlap (%):\n")
  print(round(over.mean, 2)) # Show 95%
  
  write.table(over.mean, file = file.path(sia_dir, "NicheROVER", 
              paste0("Rover_Overlap_", period, ".csv")), 
              append = F, sep = ',', row.names = T, col.names = NA)
  
# Calculate and save credible intervals
  over.cred <- apply(over.stat * 100, c(1:2, 4), quantile, prob = c(.025, .975), na.rm = TRUE)
  
write.table(over.cred, file = file.path(sia_dir, "NicheROVER", 
            paste0("Rover_Cred_", period, ".csv")), 
            append = F, sep = ',', row.names = T, col.names = NA)
  
# Niche size calculation
# Posterior distribution of niche size
  fish.size <- sapply(fish.par, function(spec) {
    apply(spec$Sigma, 3, niche.size, alpha = .95)
  })
  
write.table(fish.size, file = file.path(sia_dir, "NicheROVER", 
            paste0("Rover_Niche_Size_", period, ".csv")), 
            append = F, sep = ',', row.names = T, col.names = NA)
  
# Point estimate and standard error
  cat("Niche Size (est & se):\n")
  print(rbind(est = colMeans(fish.size), se = apply(fish.size, 2, sd)))
  
# Summary statistics for Kable
# Niche size
  size_est <- colMeans(fish.size)
  size_se <- apply(fish.size, 2, sd)
  
# Niche overlap (95%) - over.mean array is [Species A, Species B, Alpha]
  onki_on_onmy <- over.mean[1, 2, 1] # [Onki, Onmy, 95%]
  onmy_on_onki <- over.mean[2, 1, 1] # [Onmy, Onki, 95%]

# Get 95% CrIs from the 'over.cred' object
# Dim: [prob, from, to, alpha]
  onki_on_onmy_lowCI <- over.cred[1, 1, 2, 1] # prob=.025, from=Onki, to=Onmy, alpha=95%
  onki_on_onmy_highCI <- over.cred[2, 1, 2, 1] # prob=.975, from=Onki, to=Onmy, alpha=95%
  
  onmy_on_onki_lowCI <- over.cred[1, 2, 1, 1] # prob=.025, from=Onmy, to=Onki, alpha=95%
  onmy_on_onki_highCI <- over.cred[2, 2, 1, 1] # prob=.975, from=Onmy, to=Onki, alpha=95%
  
# Create a one-row data frame for this period
  period_stats <- data.frame(
    Period = period,
    Onki_Size_Est = size_est["Onki"],
    Onki_Size_SE = size_se["Onki"],
    Onmy_Size_Est = size_est["Onmy"],
    Onmy_Size_SE = size_se["Onmy"],
    Onki_on_Onmy_Overlap = onki_on_onmy,
    Onmy_on_Onki_Overlap = onmy_on_onki,
    
# Add new CrI columns
    Onki_on_Onmy_lowCI = onki_on_onmy_lowCI,
    Onki_on_Onmy_highCI = onki_on_onmy_highCI,
    Onmy_on_Onki_lowCI = onmy_on_onki_lowCI,
    Onmy_on_Onki_highCI = onmy_on_onki_highCI
  )
  
# Add to the master list
  summary_stats_list[[period]] <- period_stats
  
# Save the standard nicheROVER plots to a PDF for diagnostics
  pdf(file = file.path(sia_dir, "NicheROVER", paste0(period, "_nicheROVER_base_plots.pdf")), width = 7, height = 6)
  
# Parameter plots
  par(mar = c(4, 4, .5, .1) + .1, mfrow = c(1, 3))
  niche.par.plot(fish.par, col = clrs, plot.index = 1)
  niche.par.plot(fish.par, col = clrs, plot.index = 2)
  niche.par.plot(fish.par, col = clrs, plot.index = 1:2)
  legend("topleft", legend = names(fish.par), fill = clrs)
  
# 2D niche region plot
  par(mfrow = c(1, 1))
  fish.data <- tapply(1:nrow(nicheROVER.df), nicheROVER.df$species, function(ii) X = nicheROVER.df[ii, 2:3])
  niche.plot(niche.par = fish.par, niche.data = fish.data, pfrac = .05,
             iso.names = expression(delta^{15} * N, delta^{13} * C),
             col = clrs, xlab = expression("Isotope Ratio (\u2030)"))
  
# Overlap plot (using 95% alpha). This recalculates overlap with 1e3 nprob
  over.stat.plot <- overlap(fish.par, nreps = nsamples, nprob = 1e3, alpha = .95)
  overlap.plot(over.stat.plot, col = clrs, mean.cred.col = "black", equal.axis = TRUE,
               xlab = "Overlap Probability (%) -- Niche Region Size: 95%")
  
# Niche size boxplot
  boxplot(fish.size, col = clrs, pch = 16, cex = .5,
          ylab = "Niche Size", xlab = "Species")
  
  dev.off() #Close the PDF
 
# Prepare data for ggplot
# Use the 'over.stat.plot' object from the previous step (alpha = .95, nprob = 1e3)
  over.hist <- apply(over.stat.plot, 1:2, function(x) {
    if (any(is.na(x))) return(NULL)
    else { tmp <- x * 100 }
    tmp
  })
  
# Coho salmon (Onki) overlap on steelhead (Onmy)
  data_onki <- as.data.frame(over.hist[[1, 2]])
  colnames(data_onki) <- "Overlap"
  data_onki$Species <- "Onki"
  data_onki$Month <- period
  
# Steelhead (Onmy) overlap on coho salmon (Onki)
  data_onmy <- as.data.frame(over.hist[[2, 1]])
  colnames(data_onmy) <- "Overlap"
  data_onmy$Species <- "Onmy"
  data_onmy$Month <- period
  
# Combine and store
  combined_df <- rbind(data_onki, data_onmy)
  all_plot_data_list[[period]] <- combined_df
  
# Preparing ggplots
  # Standardized title lookup
  plot_labels <- c("Jun"="(a) Jun", "Jul"="(b) Jul", "Aug"="(c) Aug", "Sep"="(d) Sep", 
                   "Oct"="(e) Oct", "Nov"="(f) Nov", "Dec"="(g) Dec", "Overall"="(h) Overall")

  #Base plot info
  p <- ggplot(data = combined_df, aes(x = Overlap, y = after_stat(scaled), fill = Species, color = Species)) +
    # FIX: Added linewidth = 0.3 to thin the borders around the density distributions
    geom_density(alpha = 0.5, linewidth = 0.3) +
    scale_fill_manual(values = c("Onki" = "#D55E00", "Onmy" = "#0072B2")) +
    scale_color_manual(values = c("black", "black")) + #Black for contrast
    scale_x_continuous(limits = c(0, 100), breaks = c(0, 20, 40, 60, 80, 100)) +
    scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    labs(title = plot_labels[[period]]) + 
    theme_classic() +
    # FIX: Updated size = 0.5 to linewidth = 0.3 to thin the dashed/solid vertical mean lines
    geom_vline(xintercept = mean(subset(combined_df, Species == "Onki")$Overlap), linetype = "dashed", linewidth = 0.3) +
    geom_vline(xintercept = mean(subset(combined_df, Species == "Onmy")$Overlap), linetype = "solid", linewidth = 0.3) +
    theme(
      plot.title = element_text(family = "serif", hjust = 0, size = 10, face = "plain", margin = margin(b = 3, unit = "pt")),
      axis.text = element_text(family = "serif", size = 9, color = "black"),
      axis.title = element_blank(), #Removes all local axis titles
      axis.line = element_line(color = "black", linewidth = 0.3),
      axis.ticks = element_line(color = "black", linewidth = 0.3)
    )

 # Additional axes and margin logic
  top_row <- c("Jun", "Jul", "Aug", "Sep")
  bottom_row <- c("Oct", "Nov", "Dec", "Overall")
  left_col <- c("Jun", "Oct")
  right_col <- c("Sep", "Overall")
  
  # Set tight margins (2pt) for inner edges, and standard margins (5pt) for outer edges
  t_mar <- ifelse(period %in% top_row, 5, 2)
  b_mar <- ifelse(period %in% bottom_row, 12, 5)
  l_mar <- ifelse(period %in% left_col, 5, 2)
  r_mar <- ifelse(period %in% right_col, 5, 2)
  
  p <- p + theme(plot.margin = margin(t = t_mar, r = r_mar, b = b_mar, l = l_mar, unit = "pt"))

  # Remove X numbers for top row
  if(period %in% top_row) {
    p <- p + theme(axis.text.x = element_blank())
  }
  # Remove Y numbers for non-left columns
  if(!period %in% left_col) {
     p <- p + theme(axis.text.y = element_blank())
  }
  
# Legend logic (Jun Only) to match Figure 2
  if(period == "Jun") {
    p <- p + theme(
      # Pin the legend to the absolute top-left corner to avoid the density tail
      legend.position = c(0.02, 0.98),
      legend.justification = c(0, 1),
      legend.background = element_rect(fill = "white", color = "black", linewidth = 0.3),
      legend.margin = margin(2, 2, 2, 2), 
      legend.title = element_blank(),
      legend.key.size = unit(0.7, "lines"), 
      legend.text = element_text(family = "serif", size = 7), 
      text = element_text(family = "serif")
    ) +
    # Override the legend keys to have a thin black border around the fill colors
    guides(fill = guide_legend(override.aes = list(color = "black", linewidth = 0.3)))
    
  } else {
    # Hide legends for all other panels
    p <- p + theme(legend.position = "none")
  }
  
  # Save the plot object to the list
  plot_list[[period]] <- p
  
  cat(paste("--- Finished:", period, "---\n"))
}
```

### Create summary table of isotopic niche size and niche overlap

```{r kable_summary_table, echo=FALSE}
#Combine the list of data frames into one
#suppressWarnings to hide "binding factor and character..."
suppressWarnings({
   summary_table_raw <- do.call(rbind, summary_stats_list)
})

# Format results into the desired "Est (SE)" and "Overlap% (CrI)" strings
summary_table_formatted <- data.frame(
  Period = summary_table_raw$Period,
  Onki_Niche_Size = sprintf("%.2f (%.2f)", 
                          summary_table_raw$Onki_Size_Est, 
                          summary_table_raw$Onki_Size_SE),
  Onmy_Niche_Size = sprintf("%.2f (%.2f)", 
                          summary_table_raw$Onmy_Size_Est, 
                          summary_table_raw$Onmy_Size_SE),
  Onki_on_Onmy_Overlap = sprintf("%.1f%% (%.1f-%.1f)", 
                               summary_table_raw$Onki_on_Onmy_Overlap,
                               summary_table_raw$Onki_on_Onmy_lowCI,
                               summary_table_raw$Onki_on_Onmy_highCI),
  Onmy_on_Onki_Overlap = sprintf("%.1f%% (%.1f-%.1f)", 
                               summary_table_raw$Onmy_on_Onki_Overlap,
                               summary_table_raw$Onmy_on_Onki_lowCI,
                               summary_table_raw$Onmy_on_Onki_highCI)
)

#Define column names and create the kable
col_names <- c("Month", 
               "Coho (Onki)", 
               "Steelhead (Onmy)", 
               "Coho on Steelhead", 
               "Steelhead on Coho")

# Create a copy for export, apply the clean column names, and save as .csv
export_table <- summary_table_formatted
colnames(export_table) <- col_names
write.csv(export_table, file.path(sia_dir, "Tables", "NicheROVER_Overlap_Summary_Table.csv"), row.names = FALSE)

# Generate the HTML table for the document
kbl(summary_table_formatted, 
    col.names = col_names, 
    row.names = FALSE, 
    align = "lrrrr",
    caption = "Isotopic niche size (mean & SE) and 95% niche overlap (mean & 95% credible interval) estimates generated by nicheROVER.") %>%
   kable_classic(font = "Cambria", full_width = TRUE) %>%
  add_header_above(c(" " = 1, 
                   "Mean niche Size (SE)" = 2, 
                   "Mean 95% Niche Overlap (95% CI)" = 2)) %>% # Header changed
  column_spec(1, bold = TRUE) # Make the Period column bold

```

### Creating Figure 4, Probability of isotopic niche overlap

```{r Create Figure 4, fig.width=6.5, fig.height=3.97, echo=FALSE}
library(grid)
library(patchwork)

if (length(plot_list) == length(periods)) {
  
  #Arrange plots
  ordered_plots <- plot_list[c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Overall")]
  
  # Main 4 x 2 grid (just the stitched plots, no titles)
  main_grid <- wrap_plots(ordered_plots, ncol = 4, nrow = 2) &
    theme(text = element_text(family = "serif"))
  
  # Global Y canvas label
  y_expr <- expression(paste("Scaled posterior density"))
  y_grob <- textGrob(y_expr, rot = 90, gp = gpar(fontfamily = "serif", fontsize = 10))
  
  # Stitch the Y-axis directly to the grid (Keeps layout engine native)
  raw_panel <- wrap_elements(y_grob) + main_grid + plot_layout(widths = c(0.7, 25))
  
  # Add the centered X-axis title
  final_panel <- raw_panel +
    plot_annotation(
      caption = "Isotopic niche overlap probability (%)",
      theme = theme(
        # Pull text into visual center relative to the 4 panels
        plot.caption = element_text(family = "serif", size = 10, face = "plain", hjust = 0.53,
                                    # Fine-tune the vertical gap using 't' margin
                                    margin = margin(t = 2, r = 0, b = 2, l = 0, unit = "pt"))
      )
    )
  
  print(final_panel)

# Output
ggsave(file.path(sia_dir, "Figures", "Figure_4.pdf"), 
       width = 6.5, height = 3.97, units = "in", dpi = 800, 
       device = function(file, ...) quartz(file = file, type = "pdf", ...))
ggsave(file.path(sia_dir, "Figures", "Figure_4.tif"), 
       width = 6.5, height = 3.97, units = "in", 
       dpi = 800, device = "tiff", compression = "lzw")
  
} else {
  print("Skipped panel plot generation because one or more periods failed.")
}
```


Figure 4. Bayesian estimates of isotopic niche overlap (95% region) between juvenile coho salmon (Oncorhynchus kisutch; Onki) and steelhead (O. mykiss; Onmy) captured in the Scott Creek lagoon. Overlap estimates were generated for each month of the study (panels a–g) and an overall estimate (h) was produced from pooled data (all months combined). The estimates were derived from δ13C and δ15N values and represent the probability that an individual coho salmon would be found within the isotopic niche region of steelhead, and vice versa. In each panel, dashed and solid vertical lines identify the mean isotopic niche overlap (%) for coho salmon and steelhead, respectively.
